---
title: "ROCs and correlations"
format:
  html:
    embed-resources: true
    grid:
      body-width: 1800px
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  out.width = "100%"
)
library(targets)
library(patchwork)
```

::: {.panel-tabset}
## ROC curves: each dementia subtypes vs others
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(combined_roc)
```
:::

## ROC curves: each dementia subtypes vs others by sex
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
data <- tar_read(sex_specific_rocs)
data[[1]]
data[[2]]
data[[3]]
data[[4]]
```
:::

## ROC curves: each dementia subtype vs control
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(subtypes_control)
```
:::

## ROC curve: LBD vs FTD
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(roc_results)$LBD_vs_FTD[["roc_plot"]]
```
:::


## Biomarker MMSE correlations
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
mmse <- tar_read(mmse_out)
mmse[[1]]
for (marker in mmse[[2]]) {
  marker
}

```
:::

## Biomarker CDR correlations
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
cdr <- tar_read(cdr_out)
cdr[[1]]
for (marker in cdr[[2]]) {
  marker
}
```
:::

## Biomarker and CSF correlations
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(csf_out)
```
:::

## Biomarkers and PET measures correlations
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
pet <- tar_read(pet_ptau)
pet[[1]]
for (marker in pet[[2]]) {
  marker
}
```
:::

## Variable importance
::: {.my-table-wrapper style="width:60%; margin:auto;"}
#### * indicates p < 0.05
```{r}
#| results: asis
vimp <- tar_read(vimp_full)
vimp[[1]]
vimp[[2]]
vimp[[3]]
```
:::

## Variable importance by sex
::: {.my-table-wrapper style="width:60%; margin:auto;"}
#### * indicates p < 0.05
:::: {.columns}
::: {.column width="50%" padding="10%"}
```{r}
#| results: asis
vimp_f <- tar_read(vimp_females)
vimp_f[[1]]
vimp_f[[2]]
vimp_f[[3]]
```
:::

::: {.column width="50%"}
```{r}
#| results: asis
vimp_m <- tar_read(vimp_males)
vimp_m[[1]]
vimp_m[[2]]
vimp_m[[3]]
```
:::
::::
:::

```{r}
#| include: false
plots <- tar_read(subset_plots)
```

## Biomarker subset path
```{r, fig.align="center", out.width="80%"}
plots$ad
plots$ftd
plots$lbd
```

```{r}
#| include: false
plots_men <- tar_read(subset_plots_men)
plots_women <- tar_read(subset_plots_women)
```

## Biomarker subset path by sex
```{r, fig.align="center", out.width="100%"}
plots_men$ad + plots_women$ad
plots_men$ftd + plots_women$ftd
plots_men$lbd + plots_women$lbd
```
:::
