---
title: "ROCs and correlations"
format:
  html:
    embed-resources: true
    grid:
      body-width: 1800px
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  out.width = "100%"
)
```

::: {.panel-tabset}
## Subtypes vs Other Dem
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(combined_roc)
```
:::

## Subtypes by sex
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(sex_specific_rocs)
```
:::

## Subtypes vs Control
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(subtypes_control)
```
:::

## LBD vs FTD
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(roc_results)$LBD_vs_FTD[["roc_plot"]]
```
:::


## MMSE
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(mmse_out)
```
:::

## CDR
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
cdr_list <- list(
  "All" = as.data.frame(whole_inflam$cdr),
  "Any Dem" = as.data.frame(Any_inflam$cdr),
  "AD" = as.data.frame(AD_inflam$cdr),
  "LBD" = as.data.frame(LBD_inflam$cdr),
  "FTD" = as.data.frame(FTD_inflam$cdr)
)

cdr <- bind_cols(cdr_list) |>
  setNames(names(cdr_list)) |>
  rownames_to_column("Marker") |>
  relocate(Marker) |>
  mutate(Marker = case_when(
                            Marker == "mean_elisa" ~ "CD14",
                            Marker == "mean_nfl" ~ "NfL",
                            Marker == "mean_ykl" ~ "YKL-40",
                            Marker == "mean_gfap" ~ "GFAP",
                            Marker == "mean_ab42_ab40_ratio" ~ "AB42/AB40 Ratio",
                            Marker == "mean_ab40" ~ "AB40",
                            Marker == "mean_ab42" ~ "AB42",
                            Marker == "mean_tdp" ~ "TDP-43",
                            Marker == "mean_ptau181" ~ "pTau-181",
                            Marker == "mean_ptau217" ~ "pTau-217",
                            TRUE ~ Marker
                            ))
knitr::kable(cdr, digits = 3, caption = "Inflamatory Markers and CDR (age/sex adjusted)")

for (marker in markers) {
  show(df |>
    rename(
           "CD14" = "mean_elisa",
           "NfL" = "mean_nfl",
           "YKL-40" = "mean_ykl",
           "GFAP" = "mean_gfap",
           "AB42/AB40 Ratio" = "mean_ab42_ab40_ratio",
           "AB40" = "mean_ab40",
           "AB42" = "mean_ab42",
           "TDP-43" = "mean_tdp",
           "pTau-181" = "mean_ptau181",
           "pTau-217" = "mean_ptau217"
           ) |>
    filter(!is.na(get(marker)) & !is.na(cdr)) |>
    filter(Site == "Washington") |>
    ggplot(aes(x = get(marker), y = cdr, color = Diagnosis_combined)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(color = "Diagnosis") +
    xlab(marker) +
    ggtitle(paste0(marker, " vs CDR (not adjusted)")))
}
```
:::

## CSF
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
tar_read(csf_out)
```
:::

## PET / PTAU217
::: {.my-table-wrapper style="width:70%; margin:auto;"}
```{r}
whole_pet <- pet_corr(df)
Any_pet <- pet_corr(df, diagnosis = c("Alzheimer's", "Frontotermporal", "Lewy bodies"))
AD_pet <- pet_corr(df, diagnosis = c("Alzheimer's"))
FTD_pet <- pet_corr(df, diagnosis = c("Frontotermporal"))
LBD_pet <- pet_corr(df, diagnosis = c("Lewy bodies"))

pet <- bind_cols(
          as.data.frame(whole_pet$coefs),
          as.data.frame(Any_pet$coefs),
          as.data.frame(AD_pet$coefs),
          as.data.frame(FTD_pet$coefs),
          as.data.frame(LBD_pet$coefs)
          ) |>
  setNames(c(
             paste0("All (n = ", whole_pet$n, ")"),
             paste0("Any Dem (n = ", Any_pet$n, ")"),
             paste0("AD (n = ", AD_pet$n, ")"),
             paste0("LBD (n = ", FTD_pet$n, ")"),
             paste0("FTD (n = ", LBD_pet$n, ")")
          )) |>
  rownames_to_column("Measure") |>
  mutate(Measure = case_when(
                            Measure == "zscore" ~ "Z-score",
                            Measure == "raw_suvr" ~ "Raw SUVR",
                            Measure == "centiloid" ~ "Centiloid",
                            TRUE ~ Measure
                            )) |>
  relocate(Measure)

knitr::kable(pet, digits = 3, caption = "PET Measures and pTau-217 (age/sex adjusted)")

df |>
  rename(
    "raw_suvr" = "av45/PIB_fsuvr_rsf_tot_cortmean",
    "centiloid" = "Centiloid_AV45_fSUVR_TOT_CORTMEA",
  ) |>
  filter(!is.na(mean_ptau217) & !is.na(raw_suvr)) |>
  filter(Site == "Washington") |>
  ggplot(aes(x = mean_ptau217, y = raw_suvr, color = Diagnosis_combined)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    color = "Diagnosis",
    title = "Raw SUVR AB-Ratio vs pTau-217 (not age/sex adjusted)",
    x = "Mean pTau-217",
    y = "SUVR"
  )

df |>
  rename(
    "raw_suvr" = "av45/PIB_fsuvr_rsf_tot_cortmean",
    "centiloid" = "Centiloid_AV45_fSUVR_TOT_CORTMEA",
  ) |>
  filter(!is.na(mean_ptau217) & !is.na(centiloid)) |>
  filter(Site == "Washington") |>
  ggplot(aes(x = mean_ptau217, y = centiloid, color = Diagnosis_combined)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    color = "Diagnosis",
    title = "Centiloid AB-Ratio vs pTau-217 (not age/sex adjusted)",
    x = "Mean pTau-217",
    y = "Centiloid"
  )
```
:::

## VImp Full
::: {.my-table-wrapper style="width:60%; margin:auto;"}
#### * indicates p < 0.05
```{r}
#| results: asis
dem_names <- c(
               "Alzheimer's",
               "Frontotermporal",
               "Lewy bodies"
)

vimp_data <- read_data() |>
  filter(Diagnosis_combined != "Control") |>
  select(
         Diagnosis_combined,
         age_combined,
         mean_elisa,
         mean_nfl,
         mean_ykl,
         mean_gfap,
         mean_ab40,
         mean_ab42,
         mean_tdp,
         mean_ptau181,
         mean_ptau217
         ) |>
  drop_na()

vimp_full <- lapply(dem_names,
                   function(outcome) {
                     vimp_function_par(data = vimp_data,
                                       outcome_subtype = outcome)$mat |>
                     select(c(s, est, test))

                   }) |>
  setNames(dem_names)

preds <- vimp_data |>
  select(-Diagnosis_combined) |>
  names()

for (dem_name in dem_names) {
  show(kable(
        cbind(preds[as.numeric(vimp_full[[dem_name]]$s)],
              vimp_full[[dem_name]]$est,
              vimp_full[[dem_name]]$test) |>
        as_tibble() |>
        setNames(c(
                   "var_name",
                   "imp",
                   "test"
                   )) |>
        mutate(imp = round(as.numeric(imp), 3)) |>
        mutate(
          imp = paste0(imp, ifelse(test, "*", ""))
        ) |>
        select(-test) |>
        mutate(var_name = case_when(
                              var_name == "mean_elisa" ~ "CD14",
                              var_name == "mean_nfl" ~ "NfL",
                              var_name == "mean_ykl" ~ "YKL-40",
                              var_name == "mean_gfap" ~ "GFAP",
                              var_name == "mean_ab42_ab40_ratio" ~ "AB42/AB40 Ratio",
                              var_name == "mean_ab40" ~ "AB40",
                              var_name == "mean_ab42" ~ "AB42",
                              var_name == "mean_tdp" ~ "TDP-43",
                              var_name == "mean_ptau181" ~ "pTau-181",
                              var_name == "mean_ptau217" ~ "pTau-217",
                              TRUE ~ var_name
                              )),
        col.names = c("Predictor", "Importance"),
        caption = paste0("VImp for ",
                         dem_name,
                         " vs Other Dem",
                         " (n = ",
                         nrow(filter(vimp_data, Diagnosis_combined == dem_name)),
                         ")")
  ))
}
```
:::

## VImp By Sex
::: {.my-table-wrapper style="width:60%; margin:auto;"}
#### * indicates p < 0.05
:::: {.columns}
::: {.column width="50%" padding="10%"}
```{r}
#| results: asis
dem_names <- c(
               "Alzheimer's",
               "Frontotermporal",
               "Lewy bodies"
)

vimp_data_men <- read_data() |>
  filter(female != 1 & Diagnosis_combined != "Control") |>
  select(
         Diagnosis_combined,
         age_combined,
         mean_elisa,
         mean_nfl,
         mean_ykl,
         mean_gfap,
         mean_ab40,
         mean_ab42,
         mean_tdp,
         mean_ptau181,
         mean_ptau217
         ) |>
  drop_na()

vimp_men <- lapply(dem_names,
                   function(outcome) {
                     vimp_function_par(data = vimp_data_men,
                                       outcome_subtype = outcome,
                                       stratification = "men")$mat |>
                     select(c(s, est, test))

                   }) |>
  setNames(dem_names)

preds <- vimp_data_men |>
  select(-Diagnosis_combined) |>
  names()

for (dem_name in dem_names) {
  show(kable(
        cbind(preds[as.numeric(vimp_men[[dem_name]]$s)],
              vimp_men[[dem_name]]$est,
              vimp_men[[dem_name]]$test) |>
        as_tibble() |>
        setNames(c(
                   "var_name",
                   "imp",
                   "test"
                   )) |>
        mutate(imp = round(as.numeric(imp), 3)) |>
        mutate(
          imp = paste0(imp, ifelse(test, "*", ""))
        ) |>
        select(-test) |>
        mutate(var_name = case_when(
                              var_name == "mean_elisa" ~ "CD14",
                              var_name == "mean_nfl" ~ "NfL",
                              var_name == "mean_ykl" ~ "YKL-40",
                              var_name == "mean_gfap" ~ "GFAP",
                              var_name == "mean_ab42_ab40_ratio" ~ "AB42/AB40 Ratio",
                              var_name == "mean_ab40" ~ "AB40",
                              var_name == "mean_ab42" ~ "AB42",
                              var_name == "mean_tdp" ~ "TDP-43",
                              var_name == "mean_ptau181" ~ "pTau-181",
                              var_name == "mean_ptau217" ~ "pTau-217",
                              TRUE ~ var_name
                              )),
        col.names = c("Predictor", "Importance"),
        caption = paste0("VImp for ",
                         dem_name,
                         " vs Other Dem in Men",
                         " (n = ",
                         nrow(filter(vimp_data_men, Diagnosis_combined == dem_name)),
                         ")")
  ))
}
```
:::

::: {.column width="50%"}
```{r}
#| results: asis
vimp_data_women <- read_data() |>
  filter(female == 1 & Diagnosis_combined != "Control") |>
  select(
         Diagnosis_combined,
         age_combined,
         mean_elisa,
         mean_nfl,
         mean_ykl,
         mean_gfap,
         mean_ab40,
         mean_ab42,
         mean_tdp,
         mean_ptau181,
         mean_ptau217
         ) |>
  drop_na()

vimp_women <- lapply(dem_names,
                  function(outcome) {
                    vimp_function_par(data = vimp_data_women,
                                      outcome_subtype = outcome,
                                      stratification = "women")$mat |>
                    select(c(s, est, test))
                  }) |>
    setNames(dem_names)

for (dem_name in dem_names) {
  show(kable(
        cbind(preds[as.numeric(vimp_women[[dem_name]]$s)],
              vimp_women[[dem_name]]$est,
              vimp_women[[dem_name]]$test) |>
        as_tibble() |>
        setNames(c(
                   "var_name",
                   "imp",
                   "test"
                   )) |>
        mutate(imp = round(as.numeric(imp), 3)) |>
        mutate(
          imp = paste0(imp, ifelse(test, "*", ""))
        ) |>
        select(-test) |>
        mutate(var_name = case_when(
                              var_name == "mean_elisa" ~ "CD14",
                              var_name == "mean_nfl" ~ "NfL",
                              var_name == "mean_ykl" ~ "YKL-40",
                              var_name == "mean_gfap" ~ "GFAP",
                              var_name == "mean_ab42_ab40_ratio" ~ "AB42/AB40 Ratio",
                              var_name == "mean_ab40" ~ "AB40",
                              var_name == "mean_ab42" ~ "AB42",
                              var_name == "mean_tdp" ~ "TDP-43",
                              var_name == "mean_ptau181" ~ "pTau-181",
                              var_name == "mean_ptau217" ~ "pTau-217",
                              TRUE ~ var_name
                              )),
        col.names = c("Predictor", "Importance"),
        caption = paste0("VImp for ",
                         dem_name,
                         " vs Other Dem in Women",
                         " (n = ",
                         nrow(filter(vimp_data_women, Diagnosis_combined == dem_name)),
                         ")")
  ))
}
```
:::
::::
:::

```{r}
#| include: false
df <- read_search_data()

ad_n <- nrow(df[df$Diagnosis_combined == "Alzheimer's", ])
ad_full <- read_rds("results/ad_full.rds")
ad_path <- build_path(ad_full$path, ad_full$reference_auc)
ad_plot <- plot_auc_steps(ad_path, "Alzheimer's", paste0(" - n = ", ad_n))

ftd_n <- nrow(df[df$Diagnosis_combined == "Frontotemporal", ])
ftd_full <- read_rds("results/ftd_full.rds")
ftd_path <- build_path(ftd_full$path, ftd_full$reference_auc)
ftd_plot <- plot_auc_steps(ftd_path, "Frontotemporal", paste0(" - n = ", ftd_n))

lbd_n <- nrow(df[df$Diagnosis_combined == "Lewy bodies", ])
lbd_full <- read_rds("results/lbd_full.rds")
lbd_path <- build_path(lbd_full$path, lbd_full$reference_auc)
lbd_plot <- plot_auc_steps(lbd_path, "Lewy bodies", paste0(" - n = ", lbd_n))

  men <- df |>
    filter(female == 0) |>
    select(-female)
  women <- df |>
    filter(female == 1) |>
    select(-female)

```

## Subset Full
```{r, fig.align="center", out.width="80%"}
ad_plot
ftd_plot
lbd_plot
```

```{r}
#| include: false
ad_men_n <- nrow(df[df$female == 0 & df$Diagnosis_combined == "Alzheimer's", ])
ad_men <- read_rds("results/ad_men.rds")
ad_men_path <- build_path(ad_men$path, ad_men$reference_auc)
ad_men_plot <- plot_auc_steps(ad_men_path, "Alzheimer's", paste0(" - Men, n = ", ad_men_n))

ad_women_n <- nrow(df[df$female == 1 & df$Diagnosis_combined == "Alzheimer's", ])
ad_women <- read_rds("results/ad_women.rds")
ad_women_path <- build_path(ad_women$path, ad_women$reference_auc)
ad_women_plot <- plot_auc_steps(ad_women_path, "Alzheimer's", paste0(" - Women, n = ", ad_women_n))

ftd_men_n <- nrow(df[df$female == 0 & df$Diagnosis_combined == "Frontotemporal", ])
ftd_men <- read_rds("results/ftd_men.rds")
ftd_men_path <- build_path(ftd_men$path, ftd_men$reference_auc)
ftd_men_plot <- plot_auc_steps(ftd_men_path, "Frontotemporal", paste0(" - Men, n = ", ftd_men_n))

ftd_women_n <- nrow(df[df$female == 1 & df$Diagnosis_combined == "Frontotemporal", ])
ftd_women <- read_rds("results/ftd_women.rds")
ftd_women_path <- build_path(ftd_women$path, ftd_women$reference_auc)
ftd_women_plot <- plot_auc_steps(ftd_women_path, "Frontotemporal", paste0(" - Women, n = ", ftd_women_n))

lbd_men_n <- nrow(df[df$female == 0 & df$Diagnosis_combined == "Lewy bodies", ])
lbd_men <- read_rds("results/lbd_men.rds")
lbd_men_path <- build_path(lbd_men$path, lbd_men$reference_auc)
lbd_men_plot <- plot_auc_steps(lbd_men_path, "Lewy bodies", paste0(" - Men, n = ", lbd_men_n))

lbd_women_n <- nrow(df[df$female == 1 & df$Diagnosis_combined == "Lewy bodies", ])
lbd_women <- read_rds("results/lbd_women.rds")
lbd_women_path <- build_path(lbd_women$path, lbd_women$reference_auc)
lbd_women_plot <- plot_auc_steps(lbd_women_path, "Lewy bodies", paste0(" - Women, n = ", lbd_women_n))
```

## Subset By Sex
```{r, fig.align="center", out.width="100%"}
ad_men_plot + ad_women_plot
ftd_men_plot + ftd_women_plot
lbd_men_plot + lbd_women_plot
```
:::
